QPLIB: A Library of Quadratic Programming Instances

### 2016-07-22 format.tex
add my comments to the format

The QPLIB format is designed to cope with a variety of
linear and quadratic programs. We distinguish, in particular,
linear programs (LP)
\begin{align*}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{linear programs with quadratic constraints (LPQC)}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{bound-constrained quadratic programs (BQP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & x\s{L} \leq x \leq x\s{U}, \\
\intertext{quadratic programs (QP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{and quadratic program with quadratic constraints (QPQC)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U},
\end{align*}
where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
$A$ is a rectangular $m$ by $n$ constraint matrix,
$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
are vectors of appropriate sizes and $f$ is a scalar; the notation
$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
$x^T Q\s{C}_k x$.
In addition, all or a subset of the variables $x$ may be required to be
integer, while further restrictions (such as a variable being binary)
may be imposed via explicit constraints.

@@

@@

\begin{description}
\item [{[1]}]
The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\end{description}

>>>>

The QPLIB format is designed to cope with a variety of
linear and quadratic programs. We distinguish, in particular,
linear programs (LP)
\begin{align*}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{linear programs with quadratic constraints (LPQC)}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{bound-constrained quadratic programs (BQP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & x\s{L} \leq x \leq x\s{U}, \\
\intertext{quadratic programs (QP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{and quadratic program with quadratic constraints (QPQC)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U},
\end{align*}
\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations? If a linear program has quadratic constraints, how can it then still be a linear program? I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC. I further think that one could skip QCP and consider them as part of QCQP.}
where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
$A$ is a rectangular $m$ by $n$ constraint matrix,
$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
are vectors of appropriate sizes and $f$ is a scalar; the notation
$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
$x^T Q\s{C}_k x$.
In addition, all or a subset of the variables $x$ may be required to be
integer, while further restrictions (such as a variable being binary)
may be imposed via explicit constraints.

@@

\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
would reduce flexibility. Always fix the default value for coefficients to 0
(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
default and then overwrite with the given entries, which may then be zero?)
and have no default bounds for variables. If you then also adapt the suggestion
to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
a left- and right-hand-side, then there is also no need to specify a value for infinity.
This would leave out ranged constraints - but are they really used? One could
do it like CPLEX and have an extra constraint sense for ranges and an extra
section to specify the range.}

@@

\begin{description}
\item [{[1]}]
The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\todo[inline]{\small Stefan: I agree that leaving out information on
numerical properties of the matrices and vectors (convexity, etc.) is good here.
Such properties may also depend on tolerances, etc.
As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
\end{description}
